Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/proof-reconstruction.tex
[ paper ] fixing proofs.
[ paper ] rules in distinct files.

\subsubsection{Stripping a goal.}
\label{sssec:strip-a-goal}

To prove a goal, \Metis splits the goal into
disjoint cases. This process produces a list of new subgoals, the
conjunction of these subgoals implies the goal. Then, a proof of the
goal becomes, in smaller proofs, one refutation for each subgoal.
These subgoals are introduced in the \TSTP derivation with the \strip
inference rule as we show in the following excerpt.

\begin{verbatim}
fof(goal, conjecture, ((p & r) & q)).
fof(subgoal_0, plain, p,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, p => r,
    inference(strip, [], [goal])).
fof(subgoal_2, plain, (p & r) => q,
    inference(strip, [], [goal])).
\end{verbatim}

To prove the correctness of the process mentioned above,
we need to establish some lemmas before introduce our main goal of this
part, Theorem~\ref{thm:thm-strip}.

\begin{lemma}[lem-inv-uh]
  \label{lem:uh}
  If $Γ ⊢ \fuh\ φ$ then $Γ ⊢ φ$.
\end{lemma}

\begin{sketchproof}
Use induction on the cases defined by the outcome of the
$\fuh$ function in Eq.~(\ref{eq:uh-structured}).
\begin{itemize}
	\item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
	\item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
			  argument.
\vskip 2mm
\begin{itemize}

\item Case $φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒\rm{-to-}⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)$.
\begin{equation*}
  \scalebox{0.95}{
  ($\mathcal{D}_1$)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n) ∧ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{∧-proj₁}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n)$}
    \RightLabel{by def.}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n-1)$}
    \RightLabel{lem-inv-uh,}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
  }
\end{equation*}

\begin{equation*}
  (\mathcal{D}_2)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂) ∧ \fuh(φ₁ ⇒ φ₃)$}
    \RightLabel{∧-proj₂}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{by def.}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n-1)$}
    \RightLabel{lem-inv-uh.}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
    \end{bprooftree}
\end{equation*}

Finally, using the theorem \texttt{⇒∧⇒\rm{-to-}⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒\rm{-to-}⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒\rm{-to-}⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}
\item Other cases are proved in a similar way.
\end{itemize}
\end{itemize}
\end{sketchproof}

We now define the $\fstrip$ function that yields the conjunction of
subgoals that implies the goal of the problem in the \Metis \TSTP derivations.

\begin{definition}[strip]
\label{def:strip}
\label{eq:strip}

\begin{equation}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip &(φ₁ ∧ φ₂)     ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     ~(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh (¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh (φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      ~(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         ~(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         ~(\suc~n) &= ⊥\\
\fstrip &φ             ~n &= φ
\end{array}
\end{aligned}
\end{equation}
\end{definition}

% The definition above has termination problems since it is not
% a structural recursion. Therefore, we have reformulated the equation
% above to solve such an issue following the technique
% described in Section~\ref{ssec:structural-recursion}.
% We refer to $\fstrip$ as the function after applying the process mentioned
% above here and below unless otherwise stated.

\begin{lemma}[lem-inv-strip]
If $Γ ⊢ \fstrip~φ$ then $Γ ⊢ φ$.
\label{lem:lem-inv-strip}
\end{lemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula~$φ$ by following the cases in Eq.~(\ref{eq:strip}).
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.

\begin{itemize}
\item Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip~(¬¬φ₁)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ∧ φ₂$. We get a proof for each conjunct and using the
introduction rule for conjunction connective, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ φ₁$}
    \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip~(φ)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂~n$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ⇒ φ₂$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~(φ)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \fstrip~φ₂~n$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ ¬ (φ₁ ∨ φ₂)$. We show that $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$.
The result follows by introducing the conjunction and applying
De Morgan Laws.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip~φ, n)$}
\RightLabel{by def.}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip~φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁)$}
\RightLabel{lem-inv-uh,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{sketchproof}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~φ,n-1)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip~φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁ ⇒ \fstrip~φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip~φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}

\begin{lemma}[subst]
  \label{lem:subst}
  Substitution theorem.
\begin{equation*}
  \label{eq:substitution-theorem}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ$}   \AxiomC{$ψ ≡ φ$}
  \RightLabel{subst.}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\end{lemma}

We can now formulate the result that justifies the stripping strategy
of \Metis to prove goals.

\begin{theorem}[thm-strip]
\label{thm:thm-strip}
 If $s_i$ are the subgoals of the goal $φ$ for $i=1,\cdots,n$ and
 $\{ Γ ⊢ s_i \}_{i =1}^n$, then $Γ ⊢ φ$.
\end{theorem}

\begin{proof}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\left\{ Γ ⊢ s_i \right\}_{i=1}^n$}
  \RightLabel{∧-intro}
  \UnaryInfC{$Γ ⊢ \bigwedge_{i=1}^n s_i$}

  \AxiomC{$\fstrip~φ) ≡ \bigwedge_{i=1}^n s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \fstrip~φ)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

% -------------------------------------------------------------------

\subsubsection{Negating subgoals.}

Each proof of a subgoal is a refutation, thereby each proof assumes
the negation of its subgoal. The \negate rule
assumes the negation of such subgoals, and
appears after applying the \strip inference to the goal.

\begin{verbatim}
fof(goal, conjecture, (p & q)).
fof(subgoal_0, plain, (p), inference(strip, [], [goal])).
fof(subgoal_1, plain, (p => q), inference(strip, [], [goal])).
fof(negate_0_0, plain, (~ p), inference(negate, [], [subgoal_0])).
\end{verbatim}

To emulate this rule, we introduce the negation
of the subgoal by using $assume$ rule from the inferences in
Fig.~\ref{fig:CPL-inference-rules}.

% -------------------------------------------------------------------

\subsubsection{Splitting a conjunct.}
\label{sssec:splitting-a-conjunct}

The \conjunct rule extracts from a
conjunction one of its conjuncts. This rule is a generalization of the
projection rules for the conjunction connective as the
following \TSTP excerpt shows.

\begin{verbatim}
fof(normalize_1, plain, p & q & (r | ~ p),
  inference(canonicalize, [], [x])).
fof(normalize_2, plain, q,
  inference(conjunct, [], [normalize_1])).
fof(normalize_2, plain, (r | ~ p),
  inference(conjunct, [], [normalize_1])).
\end{verbatim}

\begin{definition}[conjunct]
  \label{def:conjunct}
  \begin{align*}
    \begin{split}
    &\fconjunct : \Prop \to \Prop \to \Prop\\
    &\fconjunct(φ, ψ) =
      \begin{cases}
          ψ, &\text{ if }φ ≡ ψ\\
          ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ \fconjunct(φ₁, ψ)\\
          ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ \fconjunct(φ₂, ψ)\\
          φ, &\text{ otherwise.}
        \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-conjunct]
  \label{thm:thm-conjunct}
  If $Γ ⊢ φ$ and $ψ  : \Prop$ then $Γ ⊢ \fconjunct(φ, ψ)$.
\end{theorem}
\begin{proof}\hspace{2cm}
\begin{itemize}
  \item For the case $φ ≡ ψ$, $Γ ⊢ \fconjunct(φ, ψ)$ normalizes to $Γ ⊢ ψ$.
Then, we get the desire conclusion by applying the $\fsubst$ theorem.
  \item If the proposition $φ$ is a conjunction, and we can get $ψ ≡ \fconjunct(φ_{i}, ψ)$ for some $i = 1,\ 2$, then,

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ₁ ∧ φ₂$}
  \RightLabel{∧-proj$_{i}$}
  \UnaryInfC{$Γ ⊢ φ_{i}$}
  \UnaryInfC{$Γ ⊢ \fconjunct(φ_{i}, ψ)$}
  \AxiomC{$ψ ≡ \fconjunct(φ_{i}, ψ)$}
  \RightLabel{subst.}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\item The last case follows trivially by using the same hypothesis.
\end{itemize}
\end{proof}

% subsubsection splitting-a-conjunct (end)
% -------------------------------------------------------------------

% ------------------------------------------------------------------------

\subsubsection{Reordering.}
\label{sssec:reordering}

Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows equivalence
by converting propositions to their conjunctive normal forms,
and reordering those and the inner disjunctions  to match them
In this section, we provide some theorems to perform such reordering tasks,
the goals are the following.

\begin{itemize}
  \item Reorder a disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$ to match with
$ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$. For all $i$, there exists at least one $j$ such
that $φᵢ ≡ ψⱼ$.
  \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another
CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$. For all $i$, there exists at least one $j$
such that $φᵢ ≡ ψⱼ$.
\end{itemize}

First, we define the $\rm{rdisj}$ function to convert
a disjunction in its right-associative form.
The $\rm{rconj}$ function that converts conjunction to its right-associative form
is not presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).

\begin{definition}[rdisj]
\label{def:rdisj}
  \begin{align*}
    \begin{split}
    \label{eq:rdisj-def}
      &\rm{rdisj}(φ) : \Prop \to \Prop\\
      &\rm{rdisj}(φ) =
      \begin{cases}
      \rm{rdisj}(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{ if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
      φ₁ ∨ \rm{rdisj}(φ₂),         & \text{ if } φ ≡ φ₁ ∨ φ₂\\
      φ,                      & \text{ otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-rdisj]
\label{lem:lem-rdisj}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rdisj}(φ)$.
\end{lemma}

\begin{definition}[build$_{\boldsymbol{∨}}$]
\begin{equation*}
  \begin{split}
  \label{eq:build-or}
  &\rm{build}_{∨} : \Prop → \Prop → \Prop\\
  &\rm{build}_{∨}(φ, ψ) =%\\
  % &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } \rm{build}_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation*}
\end{definition}

The $\rm{build}_{∨}$ function defined above intends to construct a
disjunction from another disjunction, specifically, this functions will
try to rearrange the disjuncts in the source formula to match with the
target disjunction formula, similar as we describe above.

\begin{lemma}[lem-build$_{\boldsymbol{∨}}$]
\label{lem:lem-build-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}(φ)$.
\end{lemma}

From now, we assume all propositions to be right-associative unless
otherwise stated.
The $\ffactor$ function in Eq.~(\ref{eq:factor-definition}) intends to
simplify a special case of disjunction, the repeated disjuncts
(e.g., $\ffactor(φ ∨ φ) \mapsto\ φ$).
Notice that other cases like $φ ∨ (ψ ∨ φ)$ do no reduce to $(ψ ∨ φ)$.

\begin{definition}[factor]
\begin{align}
\begin{split}
  \label{eq:factor-definition}
  &\ffactor : \Prop → \Prop\\
  &\ffactor(φ) =
  \begin{cases}
    φ₁,  &\text{ if }φ ≡ φ₁ ∨ φ₂ \text{, } φ₁ ≡ \ffactor(φ₂)\\
    φ,   &\text{ otherwise.}
  \end{cases}
\end{split}
\end{align}
\end{definition}

\begin{lemma}[lem-factor]
\label{lem:lem-factor}
  $Γ ⊢ φ$ then $Γ ⊢ \ffactor(φ)$.
\end{lemma}

To construct a disjunction $ψ$ from a disjunction $φ$, we have used ideas
from the description in \citeauthor{bohme2010}~\cite{bohme2010} to prove
equality between nested disjunctions.
The result is the $\rm{sbuild}_{∨}$ function
that uses every disjunct from the source formula, $φ$, to build up the
target disjunction $ψ$.

\begin{definition}[sbuild$_{\boldsymbol{∨}}$]
\begin{align*}
  \begin{split}
  \label{eq:strong-build-or-definition}
    &\rm{sbuild}_{∨} : \Prop → \Prop → \Prop\\
    &\rm{sbuild}_{∨}(φ, ψ) =%\\
    % &\hspace{3mm}
    \begin{cases}
      \ffactor(\rm{build}_{∨}(φ₁, ψ) ∨ \rm{build}_{∨}(φ₂,ψ)),
      &\text{ if }φ≡ φ₁ ∨ φ₂\\
      \rm{build}_{∨}(φ, ψ),  & \text{ otherwise.}
    \end{cases}
  \end{split}
\end{align*}
\end{definition}

\begin{lemma}[lem-sbuild$_{\boldsymbol{∨}}$]
\label{lem:lem-sbuild-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{lemma}

\begin{example}
Let see the main idea behind the lem-sbuild$_{∨}$ lemma.
We build the disjunction $((p ∨ q) ∨ r)$ from the disjunction
$(r ∨ (q ∨ p))$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ q$}
  \RightLabel{∨-intro₂}
  \UnaryInfC{$Γ ⊢ p ∨ q$}
  \RightLabel{∨-intro₁}
  \UnaryInfC{$Γ ⊢ (p ∨ q) ∨ r$}

  \AxiomC{$Γ ⊢ p$}
  \RightLabel{∨-intro₁}
  \UnaryInfC{$Γ ⊢ p ∨ q$}
  \RightLabel{∨-intro₁}
  \UnaryInfC{$Γ ⊢ (p ∨ q) ∨ r$}

  \RightLabel{∨-elim}
  \BinaryInfC{$Γ, q ∨ p ⊢ (p ∨ q) ∨ r$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ r$}
  \RightLabel{∨-intro₂}
  \UnaryInfC{$Γ ⊢ (p ∨ q) ∨ r$}

  \AxiomC{$Γ, q ∨ p ⊢ (p ∨ q) ∨ r$}

  \RightLabel{∨-elim}
  \BinaryInfC{$Γ, r ∨ (q ∨ p) ⊢ (p ∨ q) ∨ r$}

  \RightLabel{⇒-intro}
  \UnaryInfC{$Γ ⊢ r ∨ (q ∨ p) ⇒ (p ∨ q) ∨ r$}

  \end{bprooftree}
\end{equation*}
\end{example}

Notice that using $\rm{sbuild}_{∨}$ we can build not only a disjunction
with the same disjuncts of the source formula but also a complete different
formula by adding new disjuncts to the source formula.

\begin{definition}[reorder$_{\boldsymbol{∨}}$]
\begin{align*}
    \label{eq:reorder-or}
    \begin{split}
    &\freorder_{∨} : \Prop \to \Prop \to \Prop\\
    &\freorder_{∨} (φ, ψ) = \rm{sbuild}_{∨}(rdisj(φ), ψ).
    \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-reorder$_{\boldsymbol{∨}}$]
  \label{thm-reorder-or}
  If $Γ ⊢ φ$ and $ψ: \Prop$ then $Γ ⊢ \freorder_{∨}(φ, ψ)$.
\end{theorem}

\begin{proof}
Use Lemma~\ref{lem:lem-rdisj} and Lemma~\ref{lem:lem-sbuild-or}.
\end{proof}

We define similar functions and state some theorems as we previously did
for disjunctions, but instead, we treat only with conjunctive normal forms.

\begin{definition}[reorder$_{\boldsymbol{∧}}$]
  \begin{align*}
      \begin{split}
        &\freorder_{∧} : \Prop → \Prop \to \Prop\\
        &\freorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{ if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\text{, }\freorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }\freorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\\
          \fconjunct(φ, ψ), &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-reorder$_{\boldsymbol{∧}}$]
  \label{lem:lem-reorder-and}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧}(φ, ψ)$.
\end{lemma}

\begin{definition}[conjunct$_{\boldsymbol{∨}}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &\fconjunct_{∨} : \Prop → \Prop \to \Prop\\
    &\fconjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{ if }φ ≡ ψ\\
      ψ, &\text{ if }\freorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂,\ \fconjunct_{∨}(φ, ψ₁) ≡ ψ₁,\\
        &\text{and } \freorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ \fconjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ \fconjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-conjunct$_{\boldsymbol{∨}}$]
  \label{lem:lem-conjunct-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{lemma}

\begin{definition}[reorder$_{\boldsymbol{∧∨}}$]
 \begin{align*}
      \begin{split}
      &\freorder : \Prop \to \Prop \to \Prop\\
      &\freorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{ if } φ≡ψ\\
          ψ, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂,\ \freorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }\freorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂\\
          \fconjunct_{∨}(φ, ψ) &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

\begin{theorem}[thm-reorder$_{\boldsymbol{∧∨}}$]
  \label{thm:thm-reorder-and-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$.
\end{theorem}

% % --------------------------------------------------------------------

\subsubsection{Resolution.}
\label{sssec:resolve}

The \resolve rule incorporates the resolution theorem showed in
Fig.~\ref{fig:metis-inferences} for the \TSTP derivations.
This rule posses three arguments: a positive literal $l$ to apply the
resolution theorem, and two derivations.
The positive literal $l$ must occur in
the formula from the first derivation and the
negative literal must occur in the formula from the second derivation,
see the pattern of \emph{resolve} rule in Fig.~\ref{fig:metis-inferences}.

% \begin{figure}
\label{fig:resolve-tstp-example}
\begin{verbatim}
cnf(refute_0_4, plain, (~ r | p | q),
    inference(canonicalize, [], [normalize_0_5])).
cnf(refute_0_5, plain, (p | q | r),
    inference(canonicalize, [], [normalize_0_6])).
cnf(refute_0_6, plain, (p | q),
    inference(resolve, [$cnf(r)], [refute_0_5, refute_0_4])).
\end{verbatim}
% \caption{Excerpt of a \Metis \TSTP example for the \resolve rule.}
% \end{figure}

To emulate this rule, we perform a sequence of rearrangements inside the
formulas involved to match with the pattern expected by the resolution theorem
described above. We begin by defining a convenient version of
the binary resolution in Def.~\ref{def:rsol}.

\begin{definition}[rsol]
\label{def:rsol}
\begin{align*}
      \begin{split}
        &\rm{rsol} : \Prop \to \Prop\\
        &\rm{rsol}(φ) =
        \begin{cases}
          φ₂, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁, φ₂ ≡ φ₄\\
          φ₂ ∨ φ₄, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁\\
          φ, &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-rsol]
  \label{lem:lem-rsol}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rsol}(φ)$.
\end{lemma}

In Def.~\ref{def:resolve}, the proposed rule for $resolve$
has four arguments. The first three were described
above, the last one, ψ, represents the expected conclusion in the \Metis
derivation.

\begin{definition}[resolve]
\label{def:resolve}
 \begin{align*}
 \begin{split}
    &\fresolve : \Prop \to \Prop \to \Prop \to \Prop \to \Prop\\
    &\fresolve(φ₁, φ₂, l, ψ) =
      \rm{rsol}(\freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨} (φ₂, ¬ l ∨ ψ)).
  \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-resolve]
  \label{thm:thm-resolve}
  If $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l$, and $ψ : \Prop$ then
  $Γ ⊢ \fresolve(φ₁, φ₂, l, ψ)$.
\end{theorem}
\Needspace{3\baselineskip}
\begin{proof}
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₁$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
    \vskip 1.5mm
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₂$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
  \vskip 1.5mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{lem-rsol.}
    \UnaryInfC{$Γ ⊢ \rm{rsol}(\freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ))$}
  \end{bprooftree}
  \end{equation*}
\end{proof}

>>>>

\subfile{sections/proof-reconstruction/strip.tex}
\subfile{sections/proof-reconstruction/negate.tex}
\subfile{sections/proof-reconstruction/resolve.tex}
\subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/conjunct.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}
